troubleshooting

### 2024-11-21 README.Rmd
readme and vignette updates
made suggested final changes from Bioconductor review

# mspms

@@

[![R-CMD-check](https://github.com/baynec2/mspms/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/baynec2/mspms/actions/workflows/R-CMD-check.yaml)
[![Codecov test
coverage](https://codecov.io/gh/baynec2/mspms/branch/main/graph/badge.svg)](https://app.codecov.io/gh/baynec2/mspms?branch=main)

<!-- badges: end -->

@@

The goal of mspms is provide a R Package that can be used to easily and
reproducibly analyze data resulting from the [Multiplex Substrate Profiling by
Mass Spectrometry (MSP-MS) method](https://pubmed.ncbi.nlm.nih.gov/36948708/).

Additionally, we provide a [graphical user interface powered by shiny
apps](https://gonzalezlab.shinyapps.io/mspms_shiny/) that allows for a user to
utilize the method without requiring any R coding knowledge.

@@

mspms differs from existing proteomics packages that are available in the
Bioconductor project in that it is specifically designed to analyze MSP-MS data.
This involves unique preprocessing steps and data visualizations tailored to
this specific assay

@@

-   QFeatures: management and processing of peptide quantitative features and
    sample metadata.

-   MsCoreUtils: log2 transformation, imputation (QRILC), and normalization
    (center.median).

@@

-   ggplot2: data visualizations

-   heatmaply: interactive heatmaps

-   rstatix: conduct statistics

-   ggseqlogo: iceLogo motif visualization.

-   downloadthis: gives users the ability to download data from standard .html
    report generated by mspms::generate_report().

Lastly, mspms takes advantage of iceLogos ([Colaert, N. et al. Nature Methods 6,
786-787 (2009)](http://www.ncbi.nlm.nih.gov/pubmed/19876014)) to viusalize
overrepresented amino acid motifs relative to a background set by implementing
components of the Java software in R.

@@

**Note that only a subset of functions are exported to the user.** Helper
functions can be found in the \_helper_functions.R files.

@@

#### A note on colData

@@

It is very important that this contains at least the following columns
-"quantCols","group","condition",and "time".

@@

#### PEAKS

@@

```{r}
library(dplyr)
library(mspms)

@@

# Prepare the data
peaks_prepared_data <- mspms::prepare_peaks(lfq_filepath,
    colData_filepath,
    0.3,
    n_residues = 4
)

peaks_prepared_data
```

#### Fragpipe

@@

fragpipe_prepared_data
```
#### Proteome Discoverer

@@

```{r}
plot_heatmap(mspms_tidy_data)
```

@@

```{r,fig.height = 10, fig.width = 4}
sig_cleavage_data <- log2fc_t_test_data %>%
    dplyr::filter(p.adj <= 0.05, log2fc > 3)

>>>>

# `r BiocStyle::Biocpkg("mspms")`

@@

[![R-CMD-check](https://github.com/baynec2/mspms/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/baynec2/mspms/actions/workflows/R-CMD-check.yaml)
[![Codecov test
coverage](https://codecov.io/gh/baynec2/mspms/branch/main/graph/badge.svg)](https://app.codecov.io/gh/baynec2/mspms?branch=main)
[![Bioc release status](http://www.bioconductor.org/shields/build/release/bioc/mspms.svg)](https://bioconductor.org/checkResults/release/bioc-LATEST/mspms)
[![Bioc devel status](http://www.bioconductor.org/shields/build/devel/bioc/mspms.svg)](https://bioconductor.org/checkResults/devel/bioc-LATEST/mspms)
[![Bioc downloads rank](https://bioconductor.org/shields/downloads/release/mspms.svg)](http://bioconductor.org/packages/stats/bioc/mspms/)
[![Bioc support](https://bioconductor.org/shields/posts/mspms.svg)](https://support.bioconductor.org/tag/mspms)
[![Bioc history](https://bioconductor.org/shields/years-in-bioc/mspms.svg)](https://bioconductor.org/packages/release/bioc/html/mspms.html#since)
[![Bioc last commit](https://bioconductor.org/shields/lastcommit/devel/bioc/mspms.svg)](http://bioconductor.org/checkResults/devel/bioc-LATEST/mspms/)
[![Bioc dependencies](https://bioconductor.org/shields/dependencies/release/mspms.svg)](https://bioconductor.org/packages/release/bioc/html/mspms.html#since)
<!-- badges: end -->

@@

The goal of `r BiocStyle::Biocpkg("mspms")` is provide a R Package that can be used to easily and
reproducibly analyze data resulting from the [Multiplex Substrate Profiling by
Mass Spectrometry (MSP-MS) method](https://pubmed.ncbi.nlm.nih.gov/36948708/).

Additionally, we provide a [graphical user interface powered by shiny
apps](https://github.com/baynec2/mspms-shiny) that allows for a user to
utilize the method without requiring any R coding knowledge.

@@

mspms differs from existing proteomics packages that are available in the
Bioconductor project in that it is specifically designed to analyze MSP-MS data.
This involves unique preprocessing steps and data visualizations tailored to
this specific assay.

@@

-   `r BiocStyle::Biocpkg("QFeatures")`: management and processing of peptide quantitative features and
    sample metadata.

-   `r BiocStyle::Biocpkg("MsCoreUtils")`: log2 transformation, imputation (QRILC), and normalization
    (center.median).

@@

-   *ggplot2*: data visualizations

-   *heatmaply*: interactive heatmaps

-   *rstatix*: conduct statistics

-   *ggseqlogo*: iceLogo motif visualization.

-   *downloadthis*: gives users the ability to download data from standard .html
    report generated by mspms::generate_report().

Lastly, *mspms* takes advantage of iceLogos ([Colaert, N. et al. Nature Methods 6,
786-787 (2009)](http://www.ncbi.nlm.nih.gov/pubmed/19876014)) to viusalize
overrepresented amino acid motifs relative to a background set by implementing
components of the Java software in R.

@@

**Note that only a subset of functions are exported to the user.** Helper functions can be found in the
[preprocessing_helper_functions.R](https://github.com/baynec2/mspms/blob/main/R/preprocessing_helper_functions.R),
[processing_helper_functions.R](https://github.com/baynec2/mspms/blob/main/R/processing_helper_functions.R),[icelogo_helper_functions.R](https://github.com/baynec2/mspms/blob/main/R/iceLogo_helper_fuctions.R), [statistics_helper_functions.R](https://github.com/baynec2/mspms/blob/main/R/statistics_helper_functions.R), and [iceLogo_helper_functions.R](https://github.com/baynec2/mspms/blob/main/R/iceLogo_helper_fuctions.R) files.

These functions were not intended to be directly used by the user and are not exported in an effort to make the package API cleaner and more intuitive.

@@

### A note on colData

@@

**It is very important that this contains at least the following columns: quantCols","group","condition",and "time".**

@@

### PEAKS

@@

```{r}

library(dplyr)
library(mspms)

@@

# Prepare the data
peaks_prepared_data <- mspms::prepare_peaks(lfq_filepath,
  colData_filepath,
  quality_threshold = 0.3,
  n_residues = 4
)

peaks_prepared_data

```

### Fragpipe

@@

fragpipe_prepared_data
```
### Proteome Discoverer

@@

```{r,eval = FALSE}
plot_heatmap(mspms_tidy_data)
```

note that the interactive plot is not returned in this vignette 
due to size constraints.

@@
